---
title: "stats_3355_Hw2"
author: "Udit Lal"
date: "2024-02-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Problem 1

(a)
mat <- matrix(c(34, 23, 53, 6, 78, 93, 12, 41, 99), nrow = 3)
df <- as.data.frame(mat)
colnames(df) <- c("score_given_to_car_on_driving_test", "score_given_to_van_on_driving_test",  "score_given_to_truck_on_driving_test")

(b)
library(ggplot2)
head(mpg)
second_version_of_mpg <- mpg[mpg$cyl == 6, ]
second_version_of_mpg$class <- as.character(second_version_of_mpg$class)

#Problem 2

Read the .csv file
```{r}
a <- read.csv("1976-2020-senate.csv")
```

Name the variables
```{r}
names(a) <- c("Year", "State", "State_po", "State_fips", "State_cen", "State_ic", "Office", "District", "Stage", "Special", "Candidate", "Party_detailed", "Writein", "Mode", "CandidateVotes", "TotalVotes", "Unofficial", "Version",  "Party_simplified")
```

(a) Factor the variables
```{r}
a$Year <- factor(a$Year)
a$State <- factor(a$State)
a$Party_simplified <- factor(a$Party_simplified)
```

(b) Extract Texas data
```{r}
tx_data <- subset(a, State_po == "TX")
```

(c) Calculate the average and median number of voters
```{r}
avg_median_votes <- aggregate(CandidateVotes ~ Party_simplified, data = tx_data, FUN = function(x) c(avg_votes = round(mean(x), 0), median_votes = median(x)))


print(avg_median_votes)
```

(d) Identify the years in which the democratic candidate from Texas won.
```{r}
democratic_wins_tx <- subset(a, State_po == "TX" & Party_simplified == "DEMOCRAT")
max_votes <- tapply(democratic_wins_tx$CandidateVotes, democratic_wins_tx$Year, max)
democratic_wins_tx <- democratic_wins_tx[democratic_wins_tx$CandidateVotes %in% max_votes,]

print(democratic_wins_tx)
```


#Problem 3
```{r}
b <- read.csv("tae.data")
colnames(b) <- c("NativeEnglishSpeaker", "CourseInstructor", "Course", "Semester", "ClassSize", "ClassAttribute")
b$TAID <- seq(1, nrow(b))
```

(a) First variable into boolean
```{r}
tmp <- which(b$NativeEnglishSpeaker == 1)
b$NativeEnglishSpeaker[tmp] <- TRUE

tmp <- which(b$NativeEnglishSpeaker == 2)
b$NativeEnglishSpeaker[tmp] <- FALSE

b$NativeEnglishSpeaker <- as.logical(b$NativeEnglishSpeaker)
```

(b) Fourth Variable into boolean
```{r}
tmp <- which(b$Semester == 1)
b$Semester[tmp] <- FALSE

b$Semester <- as.logical(b$Semester)

tmp <- which(b$Semester == 2)
b$Semester[tmp] <- TRUE

b$Semester <- as.logical(b$Semester)
```

(c) Last variable into ordered factor
```{r}
tmp <- which(b$ClassAttribute == 1)
b$ClassAttribute[tmp] <- "Low"

tmp <- which(b$ClassAttribute == 2)
b$ClassAttribute[tmp] <- "Medium"

tmp <- which(b$ClassAttribute == 3)
b$ClassAttribute[tmp] <- "High"
```

(d) Average and Median class size
```{r}
avg_median_size <- aggregate(ClassSize ~ Semester, data = b, FUN = function(x) c(avg_size = round(mean(x), 0), median_size = median(x)))

print(avg_median_size)
```

(e) English speaking TAs
```{r}
native_regular <- sum(b$NativeEnglishSpeaker[b$Semester == TRUE])
non_native_regular <- sum(!b$NativeEnglishSpeaker[b$Semester == TRUE])

native_summer <- sum(b$NativeEnglishSpeaker[b$Semester == FALSE])
non_native_summer <- sum(!b$NativeEnglishSpeaker[b$Semester == FALSE])

print(native_regular)
print(non_native_regular)
print(native_summer)
print(non_native_summer)
```

(f) Proportion of scores for TAs by language
```{r}
English_TAs <- mean(b$ClassAttribute[b$NativeEnglishSpeaker] == "high")
Non_English_TAs <- mean(b$ClassAttribute[!b$NativeEnglishSpeaker] == "high")

print(English_TAs)
print(Non_English_TAs)
```


#Problem 4

In group projects in school there is always a difference in balance when it comes to doing work. I am usually the coordinator of the group which means that most of the time I am the one responsible for keeping others on task and distributing the workload. I always try to do my part, but it isn't always possible. In the handout it describes people from all ends of the group mates spectrum. I however have found different ways to ensure that everyone is pulling their own weight, that augmented with some of the strategies found in the article I believe that we can have a successful working group for this project. 