---
title: "stats_3355_Hw1"
author: "Udit Lal"
date: "2024-02-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

#Problem 1

```{r}
a <- 8 + 9 - 7 - (3 ** .3)
a = round(a, digits = 2)
print(a)
```

```{r}
b <- log2(sqrt((15 + 16) / (14 + 12)))
b = round(b, digits = 2)
print(b)
```
```{r}
c <- ((11 + sin(pi / 4) / (factorial(3) + abs(-10)))) ** 2
c = round(c, digits = 2)
print(c)
```
```{r}
d <- 6 + 5 - 4 / (3 ** 2)
d = round(d, digits = 2)
print(d)
```
```{r}
e <- exp(sqrt((14 + 13) / (12 + 11)))
e = round(e, digits = 2)
print(e)
```
```{r}
f <- ((11 + factorial(12) / (factorial(13) + 14))) ** 2
f = round(f, digits = 2)
print(f)
```

#Problem 2

(a) Enter this data into a data vector called RF
```{r}
RF <- c(2.60, 3.05, 3.74, 3.48, 5.49, 4.25, 2.57, 2.18, 3.14, 4.82, 3.28, 3.0)
```

(b) Name the data vector with the month abbreviation
```{r}
names(RF) <- c("January", "Febuary", "March", "April", "May", "June", "July", "August",
               "September", "October", "November", "December")
```

(c) What is the average rainfall in inches during a year (round to two decimal places)
```{r}
round(mean(RF), digits = 2)
```


(d) Find the month with the minimum rainfall and the month with maximum rainfall
```{r}
#The min is
RF[min(RF)]
#And the max is
RF[max(RF)]
```

#Problem 3

(a) Enter this data into a data vector called H2
```{r}
H2 <- c(2700, 2600, 3050, 2900, 3000, 2500, 2600, 3000, 2800, 3200, 2800, 3400)
```

(b) Name the data vector with the month abbreviation
```{r}
names(H2) <- c("January", "Febuary", "March", "April", "May", "June", "July", "August", 
               "September", "October", "November", "December")
```

(c) What is the total number of Hummer H2 sold in 2002?
```{r}
print(sum(H2))
```
(d) Find the month with the greatest increase from the previous month, and the month with the greatest decrease from the previous month
```{r}
#Greatest increase
print(which.max(diff(H2)))
#And the greatest decrease
print(which.min(diff(H2)))
```

#Problem 4

(a)
x <- c(1, -2, 3, -4, 5, 100)
y <- x * -1
y[y > 0]

(b)
# Create a sequence from 1 to 50
z <- seq(1, 50)

# Test whether each observation is even
even <- z %% 2 == 0

# Subset z to keep only the even numbers
z <- z[even]

(c)
mean <- function(x) {
sum (x) / length (x)
}

#Problem 5

```{r}
PrintSquare <- function() {
  for (i in 1:1000) {
    if (sqrt(i) %% 1 == 0) {
      print(i)
    }
  }
}

PrintSquare()

#Alternatively one could just print every number from 1 to 31^2 as sqrt(1000) is ~31.6. This the most optimal method to get the same result
#To make this an actual usable function, one could take a number as a parameter and have that be the limit of the for loop
```

#Problem 6

```{r}
isPrime <- function(n) {
  if (n <= 3) {
    return(TRUE)
  }
  
  if (n %% 2 == 0 || n %% 3 == 0) {
    return(FALSE)
  }
  
  i <- 5
  while (i * i <= n) {
    if (n %% i == 0 || n %% (i + 2) == 0) {
      return(FALSE)
    }
    i <- i + 6
  }
  
  return(TRUE)
}

twinPrime <- function(n) {
  for (i in 3:n) {
    if (isPrime(i) & isPrime(i - 2)) {
      cat('(', i, ',', (i - 2), ') ')
    }
  }
}

twinPrime(25)
```