---
title: "STATS3355_HW3"
author: "Udit Lal"
date: "2024-02-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Problem 1

step 0
```{r}
library(ggplot2)
a <- read.csv("train.csv")
```

(a) turn price_range into a factor variable
```{r}
a$price_range <- factor(a$price_range, labels = c("Low", "Medium", "High", "veryHigh"))
```

(b) Plot battery power vs ram
```{r}
ggplot(a, aes(x = battery_power, y = ram, color = price_range)) + 
  geom_point()
```


(c) Pearson Coefficient
```{r}
corr <- cor(a$ram, a$battery_power, method = 'pearson')
print(corr)
```

(d) Subset Mobile Data
```{r}
priceLow <- subset(a, price_range == "Low")
priceMedium <- subset(a, price_range == "Medium")
priceHigh <- subset(a, price_range == "High")
priceVeryhigh <- subset(a, price_range == "veryHigh")
```

(e) Pearsons for the subsets
```{r}
corrLow <- cor(priceLow$ram, priceLow$battery_power, method = 'pearson')
corrMedium <- cor(priceMedium$ram, priceMedium$battery_power, method = 'pearson')
corrHigh <- cor(priceHigh$ram, priceHigh$battery_power, method = 'pearson')
corrVeryHigh <- cor(priceVeryhigh$ram, priceVeryhigh$battery_power, method = 'pearson')

print(corrLow)
print(corrMedium)
print(corrHigh)
print(corrVeryHigh)

#these are different because High 
```

(f) trend lines
```{r}
ggplot(a, aes(x = battery_power, y = ram, color = price_range)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

(g) Average and median clock speeds
```{r}
c4 <- subset(a, n_cores == 4)
c6 <- subset(a, n_cores == 6)
c8 <- subset(a, n_cores == 8)

avgC4 <- mean(c4$clock_speed)
avgC6 <- mean(c6$clock_speed)
avgC8 <- mean(c8$clock_speed)

medC4 <- median(c4$clock_speed)
medC6 <- median(c6$clock_speed)
medC8 <- median(c8$clock_speed)

#The numbers don't change because they are uncorrelated 
```

(h) Density curves
```{r}
ggplot(a, aes(x = battery_power, y = ram, color = price_range)) +
  geom_density_2d() +
  scale_color_manual(values = c("blue", "green", "red", "purple")) +
  labs() 
```

(i) Box plots
```{r}
ggplot(a, aes(x = price_range, y = ram, fill = price_range)) +
  geom_boxplot() +
  labs()
```

(j)


(k) Factor log2(ram)
```{r}
a$logRam <- factor(round(log2(a$ram), digits = 2))
```

(l) Stacked bar plot
```{r}
ggplot(a, aes(x = logRam, fill = price_range)) +
  geom_bar(position = "stack") +
  labs(title = "Stacked Bar Plot of Price Range by log2(RAM)",
       x = "log2(RAM)",
       y = "Frequency") +
  scale_fill_manual(values = c("blue", "green", "red", "purple"), 
                    labels = c("Low", "Medium", "High", "veryHigh"))
```


#Problem 2
```{r}
b <- mpg
```

(a) Factor cylinders
```{r}
b$cyl <- factor(b$cyl, levels = c("4", "5", "6", "8"), ordered = TRUE)
```

(b) Factor transmission
```{r}
b$trans <- factor(substr(b$trans, 1, 4), levels = c("auto", "manu"))
```

(c) Factor drive
```{r}
b$drv <- factor(b$drv, levels = c("f", "r", "4"), ordered = TRUE)
```

(d) Factor fuel type
```{r}
b$fl <- factor(b$fl, levels = c("gasoline", "diesel", "other"))
```

(e) Factor class
```{r}
b$class <- factor(b$class, levels = c("2seater", "subcompact", "compact", "midsize", "suv", "minivan", "pickup"), ordered = TRUE)
```

(f) New variable country
```{r}
b$country <- NA

manuCountry <- list(
  "United States" = c("chevrolet", "dodge", "ford", "jeep", "lincoln", "mercury", "pontiac"),
  "Japan" = c("honda", "nissan", "subaru", "toyota"),
  "Germany" = c("audi", "volkswagen"),
  "South Korea" = "hyundai",
  "Great Britain" = "land rover")

for (country in names(manuCountry)) {
  b$country[b$manufacturer %in% manuCountry[[country]]] <- country
}
```

(g) Bar plot
```{r}
ggplot(b, aes(x = country, fill = country)) +
  geom_bar() +
  labs(title = "Number of Samples by Country",
       x = "Country",
       y = "Number of Samples")
```

(h) summarize US cars
```{r}
summary(b[b$country == "United States", c("displ", "cyl", "trans", "drv", "fl", "class")])
```

(i) Box plot and summary for US and Japan
```{r}
mpgUs <- (b$cty + b$hwy) / 2
mpgUs <- mpgUs[b$country == "United States"]
mpgJapan <- (b$cty + b$hwy) / 2
mpgJapan <- mpgJapan[b$country == "Japan"]

boxplot(mpgUs, mpgJapan, names = c("USA", "Japan"),
        ylab = "Combined MPG",
        main = "Combined MPG of USA and Japan Cars")

summary(mpgUs)
summary(mpgJapan)
```

(j) Histogram
```{r}
displacementUS <- b$displ[b$country == "United States"]
displacementJapan <- b$displ[b$country == "Japan"]

hist(displacementUS, main = "Engine Displacement of USA Cars", xlab = "Displacement")
hist(displacementJapan, main = "Engine Displacement of Japan Cars", xlab = "Displacement")
```

#Problem 3
(a) Team name: Swish Squad
(b) Team members names and majors: Andrew Contreras/Mathematics, Nimrod (Nico) Ohayon Rozanes/Data Science & Mathematics, Udit Lal/Economics & Computer Science
(c) Next meeting date: March 4th