---
title: "stats3355Hw4"
author: "Udit Lal"
date: "2024-02-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Problem 1
```{r}
library(ggplot2)
a <- read.csv("train.csv")
```

(a) Scatter plot
```{r}
a$price_range <- factor(a$price_range, labels = c("Low", "Medium", "High", "veryHigh"))

ggplot(a, aes(x = battery_power, y = ram, color = price_range)) + 
  geom_point()
```

(b) Trend lines
```{r}
ggplot(a, aes(x = battery_power, y = ram, color = price_range)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

(c) Density curve
```{r}
ggplot(a, aes(x = battery_power, y = ram, color = price_range)) +
  geom_density_2d() +
  scale_color_manual(values = c("blue", "green", "red", "purple")) +
  labs()

#ggplot(a, aes(y = a$ram)) +
#  geom_point() + 
#  geom_density()
```

(d) Box plot
```{r}
ggplot(a, aes(x = price_range, y = ram, fill = price_range)) +
  geom_boxplot() +
  labs()
```

(e) Violin plots
```{r}
ggplot(a, aes(x = factor(price_range), y = ram, fill = factor(price_range))) +
  geom_violin() +
  labs(x = "Price Range", y = "RAM") +
  scale_fill_manual(values = c("blue", "green", "orange", "red"), 
                    name = "Price Range") +
  ggtitle("Violin Plot of RAM by Price Range")
```

(f) Stacked bar graph
```{r}
a$logRam <- factor(round(log2(a$ram), digits = 2))

ggplot(a, aes(x = logRam, fill = price_range)) +
  geom_bar(position = "stack") +
  labs(title = "Stacked Bar Plot of Price Range by log2(RAM)",
       x = "log2(RAM)",
       y = "Frequency")
```


#Problem 2

```{r}
library(dplyr)
library(UsingR)
b <- UScereal
```

(a) Replace names
```{r}
levels(b$mfr) <- c("General Mills", "Kelloggs", "Nabisco", "Post", "Quaker Oats", "Ralston Purina")
```

(b) Factor shelf
```{r}
b$shelf <- factor(b$shelf, labels = c("low", "middle", "upper"))
```

(c) Product variable
```{r}
b$product
```
(d) Pearsons 
```{r}
corrPr <- cor(b$calories, b$protein, method = 'pearson')
corrfat <- cor(b$calories, b$fat, method = 'pearson')
corrSo <- cor(b$calories, b$sodium, method = 'pearson')
corrFi <- cor(b$calories, b$fibre, method = 'pearson')
corrCa <- cor(b$calories, b$carbo, method = 'pearson')
corrSu <- cor(b$calories, b$sugars, method = 'pearson')
corrPo <- cor(b$calories, b$potassium, method = 'pearson')

print(corrPr)
print(corrfat)
print(corrSo)
print(corrFi)
print(corrCa)
print(corrSu)
print(corrPo)
```

(e) correlation bar chart
```{r}
corr <- c(corrPr, corrfat, corrSo, corrFi, corrCa, corrSu, corrPo)
names(corr) <- c("protein", "fat", "sodium", "fibre", "carbo", "sugars", "potass")
corr <- sort(corr, decreasing = TRUE)
barplot(corr, main = "Correlation with Calories", xlab = "Nutrition Fact", ylab = "Correlation")
```
(f) Scatter plot
```{r}
largest <- names(corr)[1]

plot(b[, largest], b$calories, xlab = largest, ylab = "Calories", main = paste("Calories vs", largest))
abline(lm(b$calories ~ b[, largest]), col = "red")
```

(g) calories density curves
```{r}
ggplot(b, aes(x = calories, fill = mfr)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density Comparison of Calories by Manufacturer", x = "Calories") +
  scale_fill_manual(values = c("General Mills" = "blue", "Kelloggs" = "orange", "Others" = "grey"))
```

(h) calorie significance
```{r}
gm_k <- b[b$mfr %in% c("General Mills", "Kelloggs"), ]

ggplot(gm_k, aes(x = mfr, y = calories, fill = mfr)) +
  geom_boxplot() +
  labs(title = "Calories by Manufacturer", x = "Manufacturer", y = "Calories") 
```

(i) More box plots
```{r}
ggplot(UScereal, aes(x = mfr, y = protein)) +
  geom_boxplot(fill = "blue") +
  labs(title = "Protein Content by Manufacturer", x = "Manufacturer", y = "Protein (g)")
```

(j) Stacked bar plot
```{r}
ggplot(UScereal, aes(x = shelf, fill = mfr)) +
  geom_bar(position = "stack") +
  labs(title = "Shelf Placement by Manufacturer", x = "Shelf", fill = "Manufacturer")
```


#Problem 3
```{r}
c <- diamonds
```

(a) Histogram
```{r}
ggplot(c, aes(x = carat)) +
  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Carat", x = "Carat", y = "Frequency")
```

(b) Scatter plot price vs carat
```{r}
ggplot(diamonds, aes(x = carat, y = price, color = clarity)) +
  geom_point() +
  labs(title = "Price vs Carat by Clarity", x = "Carat", y = "Price", color = "Clarity")
```

(c) price vs carat with a smooth line
```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(title = "Price vs Carat with Smooth Line", x = "Carat", y = "Price")
```

(d) Price vs carat by clarity
```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
  geom_point() +
  facet_wrap(~ clarity) +
  labs(title = "Price vs Carat Faceted by Clarity", x = "Carat", y = "Price")
```

(f) create the charts
```{r}
ggplot(diamonds, aes(x = clarity, fill = cut)) +
  geom_bar(position = "dodge") +
  labs(title = "Count of Diamonds by Clarity and Cut", x = "Clarity", y = "Count", fill = "Cut")

ggplot(diamonds, aes(x = clarity)) +
  geom_bar(position = "dodge") +
  labs(title = "Count of Diamonds by Clarity, Separated by Cut", x = "Clarity", y = "Count") +
  facet_wrap(~ cut)
```

(g) 
```{r}

```

(h)
```{r}
ggplot(diamonds, aes(x = clarity, y = price, fill = cut)) +
  geom_boxplot() +
  labs(title = "Price vs Clarity, Segmented by Cut", x = "Clarity", y = "Price", fill = "Cut")
```

(i)
```{r}
ggplot(diamonds, aes(x = cut, y = price, fill = cut)) +
  geom_boxplot() +
  labs(title = "Price vs Cut, Segmented by Carat", x = "Cut", y = "Price", fill = "Cut")
```

Email about: 
1. Product name isnt used anywhere and doesnt have any data to be put in
2. in part (e) it referes to correlations from part (a), do you mean part (d)?